---
title: "NCAA Modeling 2017"
output:
  html_document:
    fig_height: 3
    fig_width: 5
  pdf_document:
    fig_height: 3
    fig_width: 5
---

# Commissioner's Report

```{r include=FALSE}
require(mosaic)
require(ggformula)
theme_set(theme_minimal())
require(lubridate)
require(humanparser)
require(printr)
trellis.par.set(theme = theme.mosaic())
source("Tourny.R")
source("Loaders.R")
knitr::opts_chunk$set(
  fig.align = "center", 
  comment = NA,
  echo = FALSE,
  cache = TRUE)
Bracket <- LoadBracket()
E <- LoadEntries()
G <- LoadGameScores()
T <- resultsTable(E, Bracket, G)
T$dept[T$dept == "Select a department"] <- "Math/Stat"
H <- humanparser::parse_names(sapply(E, function(x) x$name))
All <-
  H %>% 
  mutate(
    email = sapply(E, function(x) x$email),
    date  = sapply(E, function(x) as.character(x$time)),
    date2 = lubridate::ymd_hms(date) - hours(4)
  ) %>% 
  merge(T, by.x = "fullName", by.y = "name")
```

## Scores this year

The maximun score of 21 was acheived by four people this year.  


```{r echo=FALSE}
gf_histogram(~ score, data = T, binwidth = 1) # , type = "count")
```

### Winner: Amy Wilsterman

Tie breakers needed to go back to the round of 16.

Winners of the "Oh So Close Awrard" this year: John Ubels, Michelle Krul, and Steve Steenwyk.

```{r}
T %>% filter(score == 21) %>% select(name, dept, score, `score details`)
```



## How'd we do?

Here's a breakdown of scores by department.

```{r, echo = FALSE}
dotplot(dept ~ score, data = T, 
        jitter.y = TRUE, factor = 1, horizontal = TRUE, alpha = 0.5, cex = 1.5)
```

```{r, echo = FALSE}
T %>% 
  group_by(dept) %>% 
  summarise(n=n(), avg_score = round(mean(score, na.rm = TRUE), 2), max_score = max(score), min_score = min(score)) %>%
  ungroup() %>% 
  arrange(avg_score)
```

Engineering had an especially strong showing this year
with all five entries scoring at least 17 points.



## When should I submit my entry next year?

Is it better to submit as soon as the contest opens or to pour over the data for
several days? Let's see how the scores related to submission times.

```{r, echo = FALSE}
ggplot(data = All, aes(x = date2, y = score)) +
  geom_smooth(alpha = 0.5) +
  geom_point(alpha = .5) + # , position = "jitter") +
  labs(x = "time of (final) submission") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Company time?

At what time of day were people submitting their entries?  

```{r, echo = FALSE, message = FALSE}
ggplot(data = All, aes(x = hour(date2) + minute(date2)/60)) +
  geom_histogram(alpha = 0.5, binwidth = 1) +
  labs(x = "time of day of (final) submission") +
  xlim(0,24) + 
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```
