---
title: "NCAA Modeling 2016"
output: 
  pdf_document: 
    fig_height: 3
    fig_width: 5
---

# Commissioner's Report

```{r include=FALSE}
require(mosaic)
require(lubridate)
require(humanparser)
trellis.par.set(theme = theme.mosaic())
source("Tourny.R")
source("Loaders.R")
knitr::opts_chunk$set(fig.align = "center")
Bracket <- LoadBracket()
E <- LoadEntries()
G <- LoadGameScores()
T <- resultsTable(E, Bracket, G)
T$dept[T$dept == "Select a department"] <- "Math/Stat"
H <- humanparser::parse_names(sapply(E, function(x) x$name))
All <-
  H %>% 
  mutate(
    email = sapply(E, function(x) x$email),
    date  = sapply(E, function(x) as.character(x$time)),
    date2 = lubridate::ymd_hms(date) - hours(4)
  ) %>% 
  merge(T, by.x = "fullName", by.y = "name")
```

## Scores were a bit lower this year

```{r echo=FALSE}
histogram(~ score, data = T, width = 1, type = "count")
```

```{r include = FALSE}
msuPicks <- sum( sapply(E, function(x) x$teamsLogical["Michigan St"] ) )
```

This is in part due to the early loss by Michigan State.  We don't have official records going back to
check, but this was likely the single biggest waste of capital in the history of this event.
`r msuPicks` contestants chose MSU for a combined capital expenditure of 
`r msuPicks * 18` points with a return on investment of 0.

None of the contrarians were able to take advantage of this 18 point headstart -- Herb Fynewever (this year's winner) was one of those who lost MSU in the first round. 
Interestingly, the difference between this year's winning score (19) and last year's (25) is the number
of games MSU would have won had they won the tournament.  (But that would not have added 6 wins to
all `r msuPicks` contestants who chose MSU, since some of those folks also chose teams that MSU would have had to defeat along the way.)

\newpage

## How'd we do?

Here's a breakdown of scores by department.

```{r, echo = FALSE}
dotplot(dept ~ score, data = T, 
        jitter.y = TRUE, factor = 1, horizontal = TRUE, alpha = 0.5, cex = 1.5)
```

```{r, echo = FALSE}
T %>% 
  group_by(dept) %>% 
  summarise(n=n(), avg_score = round(mean(score, na.rm = TRUE), 2), max_score = max(score), min_score = min(score)) %>%
  ungroup() %>% 
  arrange(avg_score)
```


\newpage

## When should I submit my entry next year?

Is it better to submit as soon as the contest opens or to pour over the data for several days?
Let's see how the scores related to submission times.

```{r, echo = FALSE}
ggplot(data = All, aes(x = date2, y = score)) +
  geom_smooth(alpha = 0.5) +
  geom_point(alpha = .5) + # , position = "jitter") +
  labs(x = "time of (final) submission") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

## Company time?

At what time of day were people submitting their entries?  (It was Spring Break, so I guess
it really wasn't company time in quite the same way this year.)

```{r, echo = FALSE, message = FALSE}
ggplot(data = All, aes(x = hour(date2) + minute(date2)/60)) +
  geom_histogram(alpha = 0.5, binwidth = 1) +
  labs(x = "time of day of (final) submission") +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```
