---
title: "NCAA Modeling 2025"
subtitle: "Commissioner's Report"
title-block-banner: true
format:
  html: 
    theme: yeti
  typst: 
    margin: 
      x: 0.8in
brand:
  typography:
    fonts:
      - family: Bio Rhyme
        source: google
    base: Bio Rhyme
    headings: Bio Rhyme      
execute: 
  message: false
  warning: false
  fig-height: 3
  fig-width: 5
  echo: false
---


```{r setup, include = FALSE}
require(mosaic)
require(ggformula)
theme_set(theme_minimal())
theme_set(theme_bw())
require(lubridate)
library(pins)
library(gt)
library(lpSolve)
```

```{r load-data, include = FALSE, cache = TRUE}
board <-
  board_connect(server = "https://connect.cs.calvin.edu",
                key = "S2RwoRs5VIkMPCyELvj0mixL6QC4Emx1")
# require(humanparser)
require(printr)
trellis.par.set(theme = theme.mosaic())
library(madness)
# source("../Tourny.R")
# source("../Loaders.R")
knitr::opts_chunk$set(
  fig.align = "center", 
  comment = NA,
  echo = FALSE,
  cache = TRUE)

BracketM <- load_bracket("../data/2025/bracket-2025-M.csv")
BracketW <- load_bracket("../data/2025/bracket-2025-W.csv")
E <- board |> load_entries_from_pins(year = 2025)
EM <- build_entry_matrix(E, ext = "M")
EW <- build_entry_matrix(E, ext = "W")
ScoresM <- board |> pin_read("rpruim/NCAA-2025-scores-M") |>
  mutate(winner = ifelse(winner_01, away, home))
ScoresW <- board |> pin_read("rpruim/NCAA-2025-scores-W") |>
  mutate(winner = ifelse(winner_01, away, home))

TeamWinsM <- ScoresM |> group_by(winner) |> summarise(wins = n()) |>
  left_join(BracketM, by = c("winner" = "team"))
TeamWinsW <- ScoresW |> group_by(winner) |> summarise(wins = n()) |>
  left_join(BracketW, by = c("winner" = "team"))


TM <- tournament_init(
  names = BracketM[['team']], 
  seeds = BracketM[['seed']], 
  label = 'M') |>
  tournament_update(games = ScoresM[['game_number']], 
                    results = ScoresM[['winner_01']])

TW <- tournament_init(
  names = BracketW[['team']], 
  seeds = BracketW[['seed']], 
  label = 'W') |>
  tournament_update(games = ScoresW[['game_number']], 
                    results = ScoresW[['winner_01']])

ContestScoresM <- contest_scores(TM, EM)
ContestScoresW <- contest_scores(TW, EW)

library(tibble)
library(dplyr)
Results <-
  tibble(
    name = attr(ContestScoresM, "names"),
    dept = sapply(E, function(x) x$dept),
    mens = round(ContestScoresM),
    mens_raw = ContestScoresM
  ) |> full_join(
    tibble(
      name = attr(ContestScoresW, "names"),
      womens = round(ContestScoresW),
      womens_raw = ContestScoresW
    )
  ) |>
  mutate(
    difference = mens - womens,
    total = mens + womens
  )

Results2 <- Results |> filter( mens > 0, womens > 0)
ResultsW <- Results |> filter( womens > 0)
ResultsM <- Results |> filter( mens > 0)

# T$dept[T$dept == "Select a department"] <- "Math/Stat"
# H <- humanparser::parse_names(sapply(E, function(x) x$name))
# All <-
#   H |> 
#   mutate(
#     email = sapply(E, function(x) x$email),
#     date  = sapply(E, function(x) as.character(x$time)),
#     date2 = lubridate::ymd_hms(date) - hours(4)
#   ) |> 
#   merge(T, by.x = "fullName", by.y = "name")
```

# Scores this year

```{r echo=FALSE}
#| include: false
#| layout-ncol: 2
#| fig-width: 8
ResultsM |>
  gf_histogram(~ mens, binwidth = 1) 
ResultsW |>
  gf_histogram(~ womens, binwidth = 1) 
```

```{r}
Results |>
  # filter(mens > 0, womens > 0) |>
  gf_jitter(mens ~ womens, alpha = 0.9, width = 0.1, height = 0.1, 
            # label = ~ name,
            size = 2,
            color = ~ TRUE,
            # color = ~ (grepl("Pruim|Stob|Bebej|Tals|Horn|Louters", name)), 
            show.legend = FALSE) |>
  gf_abline(slope = ~ 1, intercept = ~0, color = 'steelblue', alpha = 0.6) |>
  gf_lims(x = c(-0.5, 25.5), y = c(-0.5, 25.5)) |>
  gf_refine(
    scale_color_manual(values = c('maroon', 'goldenrod'))
  )
```

# Winners 

```{r}
#| include: false
ResultsM |>
  arrange(-mens) |> 
  head(5)
ResultsW |>
  arrange(-womens) |> 
  head(5)
Results2 |>
  arrange(-total) |> 
  head(5)
ResultsM |>
  arrange(mens) |> 
  head(5)
ResultsW |>
  arrange(womens) |> 
  head(5)
```

* Stob Trophy (Men's): **Michelle Krul** (19); Runner-Up: Larry Louters (18)

* Pruim Trophy (Women's): **Eric Arnoys** (20); Runners-Up (each with 20 points): 
Mark Muyskens, Chad Tatko, Mike Stob, Gary Talsma, Dave Ross

* Comissioner's Trophy (Combined): **Chad Tatko** (36); Runner-Up: Gary Talsma (36)

* Menninga Trophy (lowest men's): **Chris Wieringa** (10); 

* Lowest Women's: **Matt Heun** (10); Runner-Up: Stan Haan (10)

Note: After tie-breakers, there were only two ties this year.

* **Don Tellinghuisen** and **Larry Louters** each had a very 
systematic score of $15 = 5 + 4 + 3 + 2 + 1 + 0$ in the women's bracket.

* **Eric Arnoys** and **David Dornbos** each had $14 = 9 + 3 + 2 + 0 + 0 + 0$ in the men's bracket.

{{< pagebreak >}}

# Men's vs Women's scores

Most of us scored better in the women's bracket.

```{r}
#| include: false
Results2 |>
  gf_histogram(~ difference) |>
  gf_labs(x = 'difference (womens - mens)')
```

```{r}
#| fig-height: 7
#| fig-width: 7
Results2 |>
  mutate( 
    name = reorder(name, - difference),
    better = ifelse(mens > womens, 'mens', 'womens')
  ) |> 
  gf_col(name ~ (- difference), fill = ~ better, # color = "gray50", 
         show.legend = FALSE) |>
  gf_labs(x = 'difference (womens - mens)') |>
  gf_refine(
    scale_fill_manual(values = c("goldenrod", "maroon"))
  )
```
{{< pagebreak >}}

# How'd we do?

Here's a breakdown of scores by department.

```{r, echo = FALSE}
#| fig-width: 8
#| fig-height: 2.8
#| layout-ncol: 1
gf_jitter(dept ~ mens, 
          data = ResultsM |> mutate(dept = reorder(dept, mens, mean)), 
          width = 0, height = 0.3, alpha = 0.8,
          shape = 16,
          size = 2, color = "maroon") |>
gf_point(
  dept ~ mens,
  data = ResultsM |> group_by(dept) |> summarise(mens = mean(mens)),
  color = "goldenrod",
  shape = 16,
  size = 2,
  alpha = 0.7
) |>
gf_point(
  dept ~ mens,
  data = ResultsM |> group_by(dept) |> summarise(mens = median(mens)),
  color = "goldenrod",
  shape = 2,
  size = 4,
  alpha = 0.7
) |>
  gf_labs(title = "Performance by Department -- Men",
          caption = "Gold circles = mean; Gold triangles = median") 
  # gf_facet_grid( ~ "Men")

  
gf_jitter(dept ~ womens, 
          data = ResultsW |> mutate(dept = reorder(dept, womens, mean)), 
          width = 0, height = 0.3, alpha = 0.8,
          shape = 16,
          size = 2, color = "maroon") |>
gf_point(
  dept ~ womens,
  data = ResultsW |> group_by(dept) |> summarise(womens = mean(womens)),
  color = "goldenrod",
  shape = 1,
  size = 4,
  alpha = 0.7
) |>
gf_point(
  dept ~ womens,
  data = ResultsW |> group_by(dept) |> summarise(womens = median(womens)),
  color = "goldenrod",
  shape = 2,
  size = 4,
  alpha = 0.7
) |>
  gf_labs(title = "Performance by Department -- Women",
          caption = "Gold circles = mean; Gold triangles = median") 
  # gf_facet_grid( ~ "Women")


gf_jitter(dept ~ total, 
          data = Results2 |> mutate(dept = reorder(dept, total, mean)), 
          width = 0, height = 0.3, alpha = 0.8,
          shape = 16,
          size = 2, color = "maroon") |>
gf_point(
  dept ~ total,
  data = Results2 |> group_by(dept) |> summarise(total = mean(total)),
  color = "goldenrod",
  shape = 1,
  size = 4,
  alpha = 0.7
) |>
gf_point(
  dept ~ total,
  data = Results2 |> group_by(dept) |> summarise(total = median(total)),
  color = "goldenrod",
  shape = 2,
  size = 4,
  alpha = 0.7
) |>
  gf_labs(title = "Performance by Department -- Total",
          caption = "Gold circles = mean; Gold triangles = median") 
  # gf_facet_grid( ~ "Total")

```

```{r, echo = FALSE, results = "asis"}
#| include: false
Results |> 
  group_by(dept) |> 
  summarise(n=n(), 
            avg = round(mean(total, na.rm = TRUE), 2), 
            max = max(total), min = min(total)) |>
  ungroup() |> 
  arrange(-avg) |>
  knitr::kable()
```

{{< pagebreak >}}

## Best possible brackets

### Men

```{r}
men_opt <-
  lp(direction = "max", all.bin = TRUE,
   objective.in = madness::dust(TeamWinsM$wins),
   const.mat = matrix(TeamWinsM$cost, nrow = 1),
   const.dir = "<=",
   const.rhs = 200
   )

TeamWinsM |> filter(men_opt$solution == 1) |>
  select(team = winner, wins, cost, seed, region ) |>
  arrange(seed) |>
  gt() |>
  gt::grand_summary_rows(c(wins, cost), fns = total ~ sum(.)) |>
  opt_table_font(
    size = 12,
    font = list(
      google_font(name = "Bio Rhyme"),
      "Cochin", "Serif")
  )
```

### Women

```{r}
women_opt <-
  lp(direction = "max", all.bin = TRUE,
   objective.in = madness::dust(TeamWinsW$wins),
   const.mat = matrix(TeamWinsW$cost, nrow = 1),
   const.dir = "<=",
   const.rhs = 200
   )
TeamWinsW |> filter(women_opt$solution == 1) |>
  select(team = winner, wins, cost, seed, region ) |>
  arrange(seed) |>
  gt() |>
  gt::grand_summary_rows(c(wins, cost), fns = total ~ sum(.)) |>
  opt_table_font(
    size = 12,
    font = list(
      google_font(name = "Bio Rhyme"),
      "Cochin", "Serif")
  )
```

```{r}
EM |> reshape2::melt() |> setNames(c("player", "team", "val")) |> 
  mutate(team_name = attr(TM, "team_names")[team]) |> 
  filter(val > 0) |>
  gf_tile(val ~ abbreviate(player, minlength = 6) + team_name) |>
  gf_theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
