---
title: "NCAA Modeling 2019"
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
---

# Commissioner's Report

```{r include=FALSE}
require(mosaic)
require(ggformula)
theme_set(theme_minimal())
require(lubridate)
require(humanparser)
require(printr)
trellis.par.set(theme = theme.mosaic())
source("../Tourny.R")
source("../Loaders.R")
knitr::opts_chunk$set(
  fig.align = "center", 
  comment = NA,
  echo = FALSE,
  cache = TRUE)
Bracket <- LoadBracket(file = "../data/bracket2019.csv")
E <- LoadEntries(path = "../data/Entries/", year = 2019)
G <- LoadGameScores(path = "../data/Scores", year = 2019)
T <- resultsTable(E, Bracket, G)
# T$dept[T$dept == "Select a department"] <- "Math/Stat"
# H <- humanparser::parse_names(sapply(E, function(x) x$name))
# All <-
#   H %>% 
#   mutate(
#     email = sapply(E, function(x) x$email),
#     date  = sapply(E, function(x) as.character(x$time)),
#     date2 = lubridate::ymd_hms(date) - hours(4)
#   ) %>% 
#   merge(T, by.x = "fullName", by.y = "name")
```

## Scores this year

```{r echo=FALSE}
gf_histogram(~ score, data = T, binwidth = 1) # , type = "count")
```

### Winner: Ryan Bebej

No tie breakers were needed this year, but Gary Talsma would
have won had the championship game result been reversed.
So he's this year's winner of the "Oh So Close Award".

### Top Ten finishers

```{r, results = "asis"}
T %>% filter(score >= 17) %>% 
  select(name, dept, score, `score details`) %>%
  knitr::kable()
```


\newpage

## How'd we do?

Here's a breakdown of scores by department.

```{r, echo = FALSE}
gf_jitter(dept ~ score, 
          data = T %>% mutate(dept = reorder(dept, score, mean)), 
          width = 0, height = 0.3, alpha = 0.8,
          shape = 1,
          size = 3, color = "maroon") %>%
gf_point(
  dept ~ score,
  data = T %>% group_by(dept) %>% summarise(score = mean(score)),
  color = "goldenrod",
  shape = 1,
  size = 4,
  alpha = 0.7
) %>%
gf_point(
  dept ~ score,
  data = T %>% group_by(dept) %>% summarise(score = median(score)),
  color = "goldenrod",
  shape = 2,
  size = 4,
  alpha = 0.7
) %>%
  gf_labs(title = "Performance by Department",
          caption = "Gold circles = mean; Gold triangles = median")
  


```

```{r, echo = FALSE, results = "asis"}
T %>% 
  group_by(dept) %>% 
  summarise(n=n(), avg_score = round(mean(score, na.rm = TRUE), 2), max_score = max(score), min_score = min(score)) %>%
  ungroup() %>% 
  arrange(-avg_score) %>%
  knitr::kable()
```

Math/Stat had an espeically strong showing this year. 
All six contestants scored better than the est entry in 
several other departments.



